---
title: "Process motif_count data"
author: "Michael Gilchrist"
date: "date: 2025-03-03"
output: pdf_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(
  warning = TRUE, # show warnings
  message = TRUE, # show messages
  error = TRUE, # do not interrupt generation in case of errors,
  echo = TRUE#,  # show R code
  ##results="asis" # Hopefully print summary() in for loops
  )


```
# Goal

- Add motif counts within songs and other information into current data structures
- Uses updated data from 2025-03-03

# Set up

## Load libraries

```{r}


## load libraries
library(stats)
#require(MASS) # provides negative binomial fitting:  glm.nb, but select conflicts with dplyr
library(tidyverse)
library(knitr)
```
## Local Functions

```{r}

kprint <- function(input, ...) {
    print(knitr::kable(input, ...))
cat('\n\n<!-- -->\n\n')
}


stringToSeconds <- Vectorize(
  function(string) {
  tmp <- str_replace(string,  "h *", " * 3600 + ") |>
    str_replace("m *", " * 60 + ") |>
    str_replace("s *", "")
  eval(parse(text=tmp))
}
)

```


```

## Load Data

```{r}

git_root <- system("git rev-parse --show-toplevel", intern=TRUE)
output_dir <- "output"
data_new <- read.csv(file.path(git_root, "data", "song", "merged.motif.data_2025-03-03.csv")) %>% tibble() 
```

## Process New Detailed Song Data

```{r}

motif_data_renamed <- data_new  |>
  rename(male = Male,
         date = Date,
         file = Clipped.File.Name,
         start_song = Timestamp.Start,
         end_song = Timestamp.End,
         motif = Number.of.Motifs,
         post_song_motif = Any.motifs.after..5.gap,
         notes = Other.notes) |>
  mutate(date = lubridate::mdy(date),
         start_song = stringToSeconds(start_song),
         end_song = stringToSeconds(end_song),
         song_start = pmin(start_song, end_song),
         song_end = pmax(start_song, end_song),        
         post_song_motif = str_detect(post_song_motif, " *(Y|y)es *"),
         song_length = song_end - song_start,
         #male = str_replace(male, "_misc_motifs", ""), ## filter later
         date_file = str_replace(file, "T[0-9]+_([^_]+)_([^_]+)_([^_]+)_[0-9]+", "\\3-\\1-\\2"),
         date_file = if_else(is.na(date_file), as.Date(NA), lubridate::ymd(date_file))
         )|>
    arrange(male, date, song_start) %>%
    ## Test to see if song_length is negative, if so, switch order of song_start and song_end
    
    #dplyr::select(!song_end) %>%  ## now redundant with info in song_length
    identity()

motif_data_detailed <- motif_data_renamed %>% filter(!(str_detect(male, "_misc_motifs")))

motif_data_misc <- motif_data_renamed %>%
    filter(str_detect(male, "_misc_motifs")) %>% 
    mutate(male = str_replace(male, "_misc_motifs", ""),
           motif_misc = motif) #filter((is.na(file) & is.na(song_start)))
## Check if there are any disagreements between date and file columns
try(
  if(
  motif_data_detailed |>  filter(date != date_file) |>
  dplyr::select(male, file, date_file, date) |>
    nrow() != 0) stop("Dates in file and date columns don't match!")
)

```

```{r}

## Extract motif data assigned to songs
## Add trial and song index columns
motif_data <- motif_data_detailed |>
    ##  filter(!(is.na(file))) |>
    arrange(male, date, song_start) %>%
    group_by(male, date) |>
    ##  mutate(trial = min_rank(date)) |>  # This doesn't work as expected because the same dates across songs are not merged, but treated as instances, so the same date is counted many times with the same ranking
    mutate(song_index = row_number()) |>
    ungroup() |>
    select(!c(start_song, end_song, date_file)) |> # start/end_song replaced by song_start/end
    relocate(c(file, post_song_motif, notes), .after=last_col()) |>
    relocate(c(song_index), .before=song_length) |>
    relocate(c(song_start, song_end), .after=song_length) |>
#    arrange(desc(song_length)) %>%
    identity()

motif_wrong <- motif_data %>% filter(song_length > 14) %>% arrange(date, file)
motif_wrong_file <- filter(data_new, Clipped.File.Name %in% motif_wrong$file)

if(length(motif_wrong)) write_csv(motif_wrong_file, file = "need.checking_2025-03-10.csv")
          
 
## Summarize song and within song motif data
motif_summary <- motif_data |>
  group_by(male, date) |>
    summarize(song_count = if_else(sum(motif) > 0, max(song_index), 0),
              song_motif_sum = sum(motif),
            song_motif_mean = mean(motif),
            song_motif_sd = sd(motif),
            song_length_sum = sum(song_length),
            song_length_sd = sd(song_length),
            song_motif_length_mean = mean(song_length/motif),
            song_motif_length_sd = sd(song_length/motif),
            post_song_sum = sum(post_song_motif),
            post_song_freq = post_song_sum/song_motif_sum
            ) %>% 
    left_join(motif_data_misc) %>% 
    mutate(motif_misc = if_else(is.na(motif_misc), 0, motif_misc)) %>%
    mutate(motif_count = song_motif_sum +  motif_misc, .before = song_count) %>%
    identity()



#%>%
#    replace(is.na(.), 0) %>% # Replace NAs with 0s
#    mutate(across(c("motif_sd", # Re-replace 0s in SD with NAs 
#                    "song_length_sd",
#                    "motif_length_sd"), 
#                  ~ifelse(.==0, NA, .)
#                  )
#           )


```

```{r}

if(interactive()) {
    file_name <- "testing.II.Rda"
} else {
    file_name <- knitr::current_input() %>% sub("\\.Rmd", ".Rda", .)
}


comment <- paste0("Data created using file data_processing_2025-03-03.Rmd on ", date()) 

save_obj <- c("motif_data", "motif_data_detailed", "motif_data_misc", "motif_summary")


## Add comments
## Note usage of eval and parse!!
for(x in save_obj) {
    my_exp <- paste0("comment(", x, ") <- \"", comment, "\"")
    parse(text = my_exp)
}

save(list = save_obj, file = paste0(file.path("output", file_name)) )

```
