---
title: "Visualize Piecewise Regression with Negative Binomial Type I Error on Real Data using `brms` Custom Family"
author: "Michael Gilchrist"
date: "Created: 2023-03-29; Compiled: `r date()`"
output: pdf_document
---

```{r setup, include=FALSE, }

knitr::opts_chunk$set(
  warning = TRUE, # show warnings
  
  message = TRUE, # show messages
  error = TRUE, # do not interrupt generation in case of errors,
  echo = TRUE,  # show R code
  fig.width = 7.5,
  fig.height = 10
)


if(interactive()) {
  default::default(.ess.eval) <- list(max.deparse.length=2E2, output = TRUE)
  output_dir <- "output"  
} else { 
  output_dir <- file.path("output", "render")
}

input_dir <- file.path("input")
```
# Goal

- Visualize two piece negative binomial type 1 formulation to data

## Recap

- Update to `2023-03-09_visualize.2023-02-28-fit...` for use with more recent fits in `2023-03-18_fit...` folder.
- Part of challenge is bimodal nature of some male's `x0` and `y0` fits
  - `T235` and `T236` in particular




## Insights


# Set up

## Install libraries

```{r, message = FALSE}

# install packages user might not have by replacing FALSE with TRUE

## load libraries
library(stats)
library(MASS) # provides negative binomial fitting:  glm.nb
library(ggplot2)
library(ggpubr)
library(grid)
library(gridExtra)
library(ggExtra)
library(cowplot)
library(GGally)
library(RColorBrewer) ## needed to have more than 8 colors with `palette="Set2"
library(broom)
library(tidyverse)
library(viridisLite)
library(cmdstanr)
library(rstan)
options(mc.cores = (parallel::detectCores()-2))
rstan_options(auto_write = TRUE)
library(brms)
library(bayesplot)
library(tidybayes)
library(loo)
library(modelr)
library(reshape2) ## provides melt

## options(ggplot2.continuous.colour="viridis",
##        ggplot2.discrete.colour="viridis",
##        ggplot2.scale_fill_discrete = scale_fill_viridis_d,
##        ggplot2.scale_fill_continuous = scale_fill_viridis_c)

library(reshape2)
library(latex2exp)

ggplot2::theme_set(theme_default(base_size = 10))
## Restore theme settings
## ggplot2::theme_set(theme_default())

n_cores <- 4

color_scheme_set("viridis")

```
## Load Local Functions


```{r}

source("../Local.Functions/local.functions_ZFI.fittings.R")

```

## Source family

```{r}

```

## Load `input`

```{r}

infile <- last(dir(file.path(input_dir, "tibbles"), "fit_tbl.*"))
## Override default fo most recent fits
infile <- "fit_tbl_adapt-delta-0.90.Rda"
infile_tbl <- file.path(input_dir, "tibbles", infile)
load(infile_tbl, verbose = TRUE)
## save filtered tibble incase it becomes corrupted locally
fit_tbl_orig <- fit_tbl
fit_tbl <- fit_tbl_orig %>% filter(!(is.na(fit))) 

```

### Add columns to `fit_tbl` for future use
```{r}

names_col <- c("obs", "epred", "pred", "plots")
fit_tbl <- add_column_safely(fit_tbl, names_col)

```

## Define and/or Expose Functions used by STAN

```{r}

source("../../../custom-brms-families/families/nbinom_type1.R")

## Define fmax() in case expose_functions()
fmax <- function(x, x0) max(x, x0)

## Create functions defined in stan's function header
## Only need to do this once since all fits use the same user
## defined functions such as "two_piece"
## Only do this if not already defined or forcing
force_expose_functions <- FALSE
if(!exists("two_piece") | force_expose_functions) expose_functions(fit_tbl[[1, "fit"]][[1]], vectorize = TRUE, show_compiler_warnings = FALSE)

```


## Generate Columns of Observed Data, Predicted Data and Expected Values 

```{r}

## add rows to fit_tbl if needed
## Can restore fit_tbl to original values using
##    fit_tbl <- fit_tbl_orig

force_generate_epred <- FALSE
force_generate_pred <- force_generate_epred
force_generate_obs <- force_generate_epred

curr_row_max <- nrow(fit_tbl)
curr_row_max <- 1

for(curr_row in 1:curr_row_max) {
  
  fit_brms <- fit_tbl[[curr_row, "fit"]][[1]]

  if(!is.null(ngrps(fit_brms))) {
    group = names(ranef(fit_brms))  
  } else {
    group = NA
  }
  
  if(length(fit_tbl[[curr_row, "epred"]][[1]]) ==0 | force_generate_epred) {
    group = names(ranef(fit_brms))
    epred <- generate_epred_data(fit_brms, group = group)
    fit_tbl[[curr_row, "epred"]] <-list(epred)
  }

}



  if(length(fit_tbl[[curr_row, "pred"]][[1]]) ==0 | force_generate_pred) {
    pred <- generate_pred_data(fit_brms)
    fit_tbl[[curr_row, "pred"]] <- list(pred)
  }

  if(length(fit_tbl[[curr_row, "obs"]][[1]]) ==0 | force_generate_obs) {
    obs <- fit_brms$data
    fit_tbl[[curr_row, "obs"]] <-list(obs)
  }

}
```

